PANCHANGE_analysis

### 2020-12-09 PANCHANGE_analysis.Rmd
added function to turn lm object into plottable dataset

# Forest Plos

Using standardised betas (2 stdev)
```{r gad standardised forest plot}
plot_models(gad_prepan_model1b_lm, std.est = "std2")

@@

>>>>

# Forest Plots

## run regression models to get normal output

###Function to create formula

```{r function create formula for lin models}

# function for formula
model1b.lm.form <- function(outcome,explan_var, data_set) {
  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c(explan_var,
                             "depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed_reordered",
                             "Ethnicity_collapsed",
                             "key_worker",
                             "time_diff_coping_lockdown_march_23"
  )
  # create the formula
  glm.formula <- paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ ")

  # assign it globally with the right variable name.
  assign(paste0(outcome,explan_var,".glmformula"),
         glm.formula,
         envir = .GlobalEnv
  )

@@

###Standardise linear variables

```{r standardise linear variables}
lin_preds <- c("time_diff_coping_lockdown_march_23",
               "gad.sum_score_prepan",
               "phq.sum_score_prepan",
               "pcl.sum_score_prepan",
               "gad.sum_score_retro",
               "phq.sum_score_retro",
               "pcl.sum_score_retro"
               )

dat.prepan[(names(dat.prepan) %in% lin_preds)] <- apply(dat.prepan[(names(dat.prepan) %in% lin_preds)],2,scale)
dat.retro[(names(dat.retro) %in% lin_preds)] <- apply(dat.retro[(names(dat.retro) %in% lin_preds)],2,scale)

## Run models

```{r create formulas for forest models}

model1b.lm.form("gad.sum_score_base", "gad.sum_score_prepan", dat.prepan)
model1b.lm.form("gad.sum_score_base", "gad.sum_score_retro", dat.retro)

model1b.lm.form("phq.sum_score_base", "phq.sum_score_prepan", dat.prepan)
model1b.lm.form("phq.sum_score_base", "phq.sum_score_retro", dat.retro)

model1b.lm.form("pcl.sum_score_base", "pcl.sum_score_prepan", dat.prepan)
model1b.lm.form("pcl.sum_score_base", "pcl.sum_score_retro", dat.retro)

```{r run models for all forest plot items}

gad.prepan.lm <-  lm(formula = gad.sum_score_base.gad.sum_score_prepan.glmformula,
                            dat.prepan)
gad.retro.lm <- lm(formula = gad.sum_score_base.gad.sum_score_retro.glmformula,
                            dat.retro)

phq.prepan.lm <-  lm(formula = phq.sum_score_base.phq.sum_score_prepan.glmformula,
                            dat.prepan)
phq.retro.lm <- lm(formula = phq.sum_score_base.phq.sum_score_retro.glmformula,
                            dat.retro)

pcl.prepan.lm <-  lm(formula = pcl.sum_score_base.pcl.sum_score_prepan.glmformula,
                            dat.prepan)
pcl.retro.lm <- lm(formula = pcl.sum_score_base.pcl.sum_score_retro.glmformula,
                            dat.retro)

###Subset retrospective data by samples

```{r subset data for forests into samples}

dat.retro.glad <- dat.retro %>%
  filter(Sample == "GLAD" | 
           Sample == "EDGI")

dat.retro.nbr <- dat.retro %>%
  filter(Sample == "NBR")

dat.retro.ramp <- dat.retro %>%
  filter(Sample == "RAMP")

```{r run models for rero by sample forest plots}

gad.retro.lm.glad <- lm(formula = gad.sum_score_base.gad.sum_score_retro.glmformula,
                            dat.retro.glad)
gad.retro.lm.nbr <- lm(formula = gad.sum_score_base.gad.sum_score_retro.glmformula,
                            dat.retro.nbr)
gad.retro.lm.ramp <- lm(formula = gad.sum_score_base.gad.sum_score_retro.glmformula,
                            dat.retro.ramp)

phq.retro.lm.glad <- lm(formula = phq.sum_score_base.phq.sum_score_retro.glmformula,
                            dat.retro.glad)
phq.retro.lm.nbr <- lm(formula = phq.sum_score_base.phq.sum_score_retro.glmformula,
                            dat.retro.nbr)
phq.retro.lm.ramp <- lm(formula = phq.sum_score_base.phq.sum_score_retro.glmformula,
                            dat.retro.ramp)

pcl.retro.lm.glad <- lm(formula = pcl.sum_score_base.pcl.sum_score_retro.glmformula,
                            dat.retro.glad)
pcl.retro.lm.nbr <- lm(formula = pcl.sum_score_base.pcl.sum_score_retro.glmformula,
                            dat.retro.nbr)
pcl.retro.lm.ramp <- lm(formula = pcl.sum_score_base.pcl.sum_score_retro.glmformula,
                            dat.retro.ramp)

## Function to create dataframes ready for forest plots from lm models. 
```{r function to create data for lm forest plots}

# function to create dataframe from model output

lmoutput_format <- function(lmobject,explan_var,dfname){

  ## Row labels
  independent.variables <- c("intercept",
                             explan_var,
                             "comorbid depression and anxiety compared to control",
                             "depressive disorder compared to control",
                             "anxiety disorder compared to control",
                             "eating disorders compared to control",
                             "obsessive compulsive disorders compared to control",
                             "comorbid psychotic and bipolar disorder compared to control",
                             "psychotic disorder compared to control",
                             "bipolar disorder compared to control",
                             "mhd.posttraumatic_stress_disorder_ptsd compared to control",
                             "autism_spectrum_disorder compared to control",
                             "mhd.attention_deficit_hyperactivity_disorder compared to control",
                             "mhd.personality_disorder compared to control",
                             "Female compared to male",
                             "Non-binary/Self-defined compared to male",
                             "16-18 compared to 26-35",
                             "19-25 compared to 26-35",
                             "36-45 compared to 26-35",
                             "46-55 compared to 26-35",
                             "56-65 compared to 26-35",
                             "66-70 compared to 26-35",
                             "71-75 compared to 26-35",
                             "76+ compared to 26-35",
                             "Mixed or multiple ethnicities compared to White",
                             "Asian compared to White",
                             "Black compared to White",
                             "Other compared to White",
                             "key_worker compared to non-keyworker",
                             "time_diff_coping_lockdown_march_23")
  ## categories to facet on  
  categories <- c("Intercept",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Gender",
                  "Gender",
                  "Age",
                  "Age",
                  "Age",
                  "Age",
                  "Age",
                  "Age",
                  "Age",
                  "Age",
                  "Ethnicity",
                  "Ethnicity",
                  "Ethnicity",
                  "Ethnicity",
                  "Other",
                  "Other")

  # create data frame
  df <- as.data.frame(coef(summary(lmobject)),row.names=F)

  # add row labels
  df$RowLabels <- independent.variables

  # Add category labels

  df$Categories <- categories

  # calculate upper and lower confidence intervals
  df <- df %>%
    mutate(LowerBound = Estimate - (1.96*`Std. Error`),
           UpperBound = Estimate + (1.96*`Std. Error`))

  assign(dfname, df, envir = .GlobalEnv)
